---
title: "Hyperspectral analysis in the Greater Cape Floristic Region of South Africa"
author: Willow Flood
date: Fall 2022 Semester
---

# Introduction 

The problem being addressed in this project is looking at the Peninsula Sandstone Fynbos ecosystem in the Southwest Fynbos Bioregion in the Greater Cape Floristic Region and how the reflectance is affected by wildfires. This will be done by comparing the time since fire data and hyperspectral reflectance data to see what differences there are at different times since the fire occurred. This will help to show how this ecosystem's reflectance changes due to a change in the composition of the ecosystem. (Need to bulk this section up to get to 200 words)  

# Materials and methods 

The data being used for this project come from a variety of sources. The hyperspectral data comes from the PRISMA Satellite. It captures the Greater Cape Floristic Region on 2022/01/22. The time since fire data is taken from the AdamWilsonLab emma_envdata repository on GitHub. The date of the time since fire data is 2022/01/17. The ecosystem data was collected from the South African National Biodiversity Institute (SANBI) 2018 National Vegetation Map Project (VEGMAP). 

These are the libraries that were used in this project. (need to edit this)
```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(raster)
library(rgdal)
library(plyr)
library(piggyback)
library(prismaread)
library(sf)
library(mapview)
library(stars)
library(FNN)
```
These are the days since fire and vegetation map data sets that were used for this project.
```{r message=FALSE, warning=FALSE}
dir.create("data", showWarnings = F)
piggyback::pb_download(
  "2022_01_17.tif",
  repo = "AdamWilsonLab/emma_envdata",
  tag = "processed_ndvi_relative_days_since_fire",
  dest = "data")
days_fire = read_stars("data/2022_01_17.tif")

piggyback::pb_download(
  repo = "geo511-2022/final_project-wmflood",
  tag = "VEGMAP",
  dest = "data")
vegmap = st_read("data/NVM2018_AEA_V22_7_16082019_final.shp")
```
This is the PRISMA satellite data used in this project.
```{r message=FALSE, warning=FALSE}
piggyback::pb_download(
  "PRS_L2D_STD_20220122090206_20220122090210_0001.he5.zip",
  repo = "geo511-2022/final_project-wmflood",
  tag = "PRISMA",
  dest = "data")

l2d_zip_path = file.path("data/PRS_L2D_STD_20220122090206_20220122090210_0001.he5.zip")
unzip(l2d_zip_path, exdir = "data")

idx_out_dir = file.path(tempdir(), "prismaread/indices")
dir.create(dirname(idx_out_dir))
pr_convert(
  in_file = "data/PRS_L2D_STD_20220122090206_20220122090210_0001.he5",
  out_format = "GTiff",
  out_folder = idx_out_dir,
  indexes = "hNDVI",
  overwrite = TRUE
)

```
Reading the PRISMA data. This step is reading the hyperspectral NDVI for the project.
```{r}
file_var = list.files(idx_out_dir, full.names = TRUE)

idx_hNDVI = read_stars(file_var[2]) 
```
Map of the Hyperspectral NDVI.
```{r echo=FALSE,warning=FALSE,message=FALSE}
mapview(idx_hNDVI, maxpixels = 154295)
```
Converting the days_fire data to the same crs as the idx_hNDVI and cropping it to the same area.
```{r warning=FALSE}
#Converting the days_fire data to the same Coordinate Reference System
days_fire_proj = st_transform(days_fire, crs = st_crs(idx_hNDVI))

#Cropping the vegmap_proj and the days_fire_proj
fire_crop = st_crop(days_fire_proj, idx_hNDVI, crop = TRUE)
```
Map of the Days Since Fire data crop.
```{r echo=FALSE,warning=FALSE,message=FALSE}
mapview(fire_crop)
```
Reprojecting the vegmap data set to match the idx_out_NDVI. Then cropping the vegmap data sets to the the same extent. Finally, converting it to an sp object to be better used with the raster objects.
```{r warning=FALSE}
#Reprojecting the vegmap data
vegmap_proj = st_transform(vegmap, crs = st_crs(idx_hNDVI))

#Cropping the vegmap data to the study area
vegmap_crop = st_crop(vegmap_proj, idx_hNDVI)

#Selecting the Peninsula Sandstone Fynbos
Fynbos_sf = vegmap_crop %>%
  filter(Name_18 == "Peninsula Sandstone Fynbos")

#Converting vegmap to an sp object
Fynbos_sp = as(Fynbos_sf, "Spatial") #may not need
```
Map of the Peninsula Fynbos Sandstone.
```{r echo=FALSE,warning=FALSE,message=FALSE}
mapview(Fynbos_sf)
```
Cropping the fire_crop and the idx_hNDVI to only the regions of the Fynbos_sf.
```{r warning=FALSE}
fire_fynbos = st_crop(fire_crop, Fynbos_sf, crop = TRUE)

hNDVI_fynbos = st_crop(idx_hNDVI, Fynbos_sf, crop = TRUE)
```
Current working plan. Spatially join the hNDVI_fynbos and fire_fynbos so that values can be graphed.


# Results 

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r}

```

# Conclusions 

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
