---
title: "Hyperspectral analysis in the Greater Cape Floristic Region of South Africa"
author: Willow Flood
date: Fall 2022 Semester
---

# Introduction

The problem being addressed in this project is looking at the Peninsula Sandstone Fynbos ecosystem in the Southwest Fynbos Bioregion in the Greater Cape Floristic Region and how the Normalized Difference Vegetation Index (NDVI) is affected by wildfires. This will be done by comparing the time since fire data and hyperspectral NDVI data to see what differences there are at different times since the fire occurred. This will help to show how this ecosystem's NDVI changes due to a change in the composition of the ecosystem after a wildfire. (Need to bulk this section up to get to 200 words)  

# Materials and methods

The data being used for this project came from a variety of sources. The hyperspectral data comes from the PRISMA Satellite project. The data captures the Greater Cape Floristic Region on 2022/01/22 since it was a relativly clear day and it is the closest date to the time since fire data. The time since fire data is taken from the AdamWilsonLab emma_envdata repository on GitHub. The date of the time since fire data is 2022/01/17. The ecosystem data was collected from the South African National Biodiversity Institute (SANBI) 2018 National Vegetation Map Project (VEGMAP). 

These are the libraries that were used in this project. (need to edit this)
```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(raster)
library(rgdal)
library(plyr)
library(piggyback)
library(prismaread)
library(sf)
library(mapview)
library(stars)
library(FNN)
```
These are the days since fire and vegetation map data sets that were used for this project. Using the package piggyback, data was collected from GitHub repositories.
```{r message=FALSE, warning=FALSE}
dir.create("data", showWarnings = F)
piggyback::pb_download(
  "2022_01_17.tif",
  repo = "AdamWilsonLab/emma_envdata",
  tag = "processed_ndvi_relative_days_since_fire",
  dest = "data")
days_fire = read_stars("data/2022_01_17.tif")

piggyback::pb_download(
  repo = "geo511-2022/final_project-wmflood",
  tag = "VEGMAP",
  dest = "data")
vegmap = st_read("data/NVM2018_AEA_V22_7_16082019_final.shp")
```
This is the PRISMA satellite data used in this project. The data was downloaded from GitHub using the piggyback method, and the data was read using the prismaread package. The raster data for this and the time since fire data is also read in using the stars package. The stars package was used instead of raster because of it's ability to interface with sf data. Since the SANBI vegetation coverage map is read in as sf, stars was the more compatible choice.
```{r message=FALSE, warning=FALSE}
piggyback::pb_download(
  "PRS_L2D_STD_20220122090206_20220122090210_0001.he5.zip",
  repo = "geo511-2022/final_project-wmflood",
  tag = "PRISMA",
  dest = "data")

l2d_zip_path = file.path("data/PRS_L2D_STD_20220122090206_20220122090210_0001.he5.zip")
unzip(l2d_zip_path, exdir = "data")

idx_out_dir = file.path(tempdir(), "prismaread/indices")
dir.create(dirname(idx_out_dir))
pr_convert(
  in_file = "data/PRS_L2D_STD_20220122090206_20220122090210_0001.he5",
  out_format = "GTiff",
  out_folder = idx_out_dir,
  indexes = "hNDVI",
  overwrite = TRUE
)
```
Reading the PRISMA data. In order to to read the PRISMA data from the idx_out_dir the variable file_var was created. This is then used to access the specific file path for the hyperspectral NDVI that was processed.
```{r}
file_var = list.files(idx_out_dir, full.names = TRUE)

idx_hNDVI = read_stars(file_var[2]) 
```
Map of the Hyperspectral NDVI.
```{r echo=FALSE,warning=FALSE,message=FALSE}
mapview(idx_hNDVI, maxpixels = 154295)
```
Reprojecting the days_fire data to the same crs as the idx_hNDVI and cropping it to the same area.
```{r warning=FALSE}
#Converting the days_fire data to the same Coordinate Reference System
days_fire_proj = st_transform(days_fire, crs = st_crs(idx_hNDVI))

#Cropping the days_fire_proj
fire_crop = st_crop(days_fire_proj, idx_hNDVI, crop = TRUE)
```
Reprojecting the vegmap data set to the same crs as the idx_out_NDVI. Then cropping the vegmap data sets to the the same extent.
```{r warning=FALSE}
#Reprojecting the vegmap data
vegmap_proj = st_transform(vegmap, crs = st_crs(idx_hNDVI))

#Cropping the vegmap data to the study area
vegmap_crop = st_crop(vegmap_proj, idx_hNDVI)

#Selecting the Peninsula Sandstone Fynbos
Fynbos_sf = vegmap_crop %>%
  filter(Name_18 == "Peninsula Sandstone Fynbos")
```
Map of the Peninsula Fynbos Sandstone.
```{r echo=FALSE,warning=FALSE,message=FALSE}
mapview(Fynbos_sf)
```
The next step is cropping the fire_crop and the idx_hNDVI to only the regions of the Fynbos_sf.
```{r warning=FALSE}
fire_fynbos = st_crop(fire_crop, Fynbos_sf, crop = TRUE)

hNDVI_fynbos = st_crop(idx_hNDVI, Fynbos_sf, crop = TRUE)
```
Because the two stars objects fire_fynbos and hNDVI_fynbos have different spatial resolutions, the two objects can't be merged. This leaves conducting a spatial join of the two objects.
```{r}
fire_fynbos_sf = st_as_sf(fire_fynbos, as_points = FALSE, na.rm = TRUE)
hndvi_fynbos_sf = st_as_sf(hNDVI_fynbos, as_points = TRUE, na.rm = TRUE)
hndvi_fire = st_join(fire_fynbos_sf, hndvi_fynbos_sf, join = st_nearest_feature, left = TRUE)
```

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r}
ggplot(hndvi_fire, aes(x=hndvi_fire$`2022_01_17.tif`, y=hndvi_fire$PRS_L2D_STD_20220122090206_20220122090210_0001_hNDVI.tif)) +
  geom_line()
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

Busetto, L., Ranghetti, L. (2020) prismaread: A tool for facilitating access and analysis of PRISMA L1/L2 hyperspectral imagery v1.0.0, URL: https://irea-cnr-mi.github.io/prismaread/, doi: https://doi.org/10.5281/zenodo.4019081
