---
title: "Hyperspectral analysis in the Greater Cape Floristic Region of South Africa"
author: Willow Flood
date: Fall 2022 Semester
---

# Introduction

The problem being addressed in this project is looking at the Peninsula Sandstone Fynbos ecosystem in the Southwest Fynbos Bioregion in the Greater Cape Floristic Region and how the Normalized Difference Vegetation Index (NDVI) is affected by wildfires. This will be done by comparing the time since fire data and hyperspectral NDVI data to see what differences there are at different times since the fire occurred. This will help to show how this ecosystem's NDVI changes due to a change in the composition of the ecosystem after a wildfire. (Need to bulk this section up to get to 200 words)  

# Materials and methods

The data being used for this project came from a variety of sources. The hyperspectral data comes from the PRISMA Satellite project. The data captures the Greater Cape Floristic Region on 2022/01/22 since it was a relatively clear day and it is the closest date to the time since fire data. The time since fire data is taken from the AdamWilsonLab emma_envdata repository on GitHub. The date of the time since fire data is 2022/01/17. The ecosystem data was collected from the South African National Biodiversity Institute (SANBI) 2018 National Vegetation Map Project (VEGMAP). 

These are the libraries that were used in this project.
```{r warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(piggyback)
library(prismaread)
library(sf)
library(mapview)
library(stars)
library(FNN)
```
These are the days since fire and vegetation map data sets that were used for this project. Using the package piggyback, data was collected from GitHub repositories.
```{r message=FALSE, warning=FALSE}
dir.create("data", showWarnings = F)
piggyback::pb_download(
  "2022_01_17.tif",
  repo = "AdamWilsonLab/emma_envdata",
  tag = "processed_ndvi_relative_days_since_fire",
  dest = "data")
days_fire = read_stars("data/2022_01_17.tif")

piggyback::pb_download(
  repo = "geo511-2022/final_project-wmflood",
  tag = "VEGMAP",
  dest = "data")
vegmap = st_read("data/NVM2018_AEA_V22_7_16082019_final.shp")
```
This is the PRISMA satellite data used in this project. The data was downloaded from GitHub using the piggyback method, and the data was read using the prismaread package. The raster data for this and the time since fire data is also read in using the stars package. The stars package was used instead of raster because of its ability to interface with sf data. Since the SANBI vegetation coverage map is read in as sf, stars was the more compatible choice.
```{r message=FALSE, warning=FALSE}
piggyback::pb_download(
  "PRS_L2D_STD_20220122090206_20220122090210_0001.he5.zip",
  repo = "geo511-2022/final_project-wmflood",
  tag = "PRISMA",
  dest = "data")

l2d_zip_path = file.path("data/PRS_L2D_STD_20220122090206_20220122090210_0001.he5.zip")
unzip(l2d_zip_path, exdir = "data")

idx_out_dir = file.path(tempdir(), "prismaread/indices")
dir.create(dirname(idx_out_dir))
pr_convert(
  in_file = "data/PRS_L2D_STD_20220122090206_20220122090210_0001.he5",
  out_format = "GTiff",
  out_folder = idx_out_dir,
  indexes = "hNDVI",
  overwrite = TRUE
)
```
Reading the PRISMA data. In order to read the PRISMA data from the idx_out_dir the variable file_var was created. This is then used to access the specified file path for the hyperspectral NDVI that was processed.
```{r}
file_var = list.files(idx_out_dir, full.names = TRUE)

idx_hNDVI = read_stars(file_var[2]) 
```
Map of the Hyperspectral NDVI.
```{r echo=FALSE,warning=FALSE,message=FALSE}
mapview(idx_hNDVI, maxpixels = 154295)
```
Reprojecting the days_fire data to the same crs as the idx_hNDVI and cropping it to the same area.
```{r warning=FALSE}
#Converting the days_fire data to the same Coordinate Reference System
days_fire_proj = st_transform(days_fire, crs = st_crs(idx_hNDVI))

#Cropping the days_fire_proj
fire_crop = st_crop(days_fire_proj, idx_hNDVI, crop = TRUE)
```
Reprojecting the vegmap data set to the same crs as the idx_out_NDVI. Then cropping the vegmap data sets to the same extent.
```{r warning=FALSE}
#Reprojecting the vegmap data
vegmap_proj = st_transform(vegmap, crs = st_crs(idx_hNDVI))

#Cropping the vegmap data to the study area
vegmap_crop = st_crop(vegmap_proj, idx_hNDVI)

#Selecting the Peninsula Sandstone Fynbos
Fynbos_sf = vegmap_crop %>%
  filter(Name_18 == "Peninsula Sandstone Fynbos")
```
Map of the Peninsula Fynbos Sandstone.
```{r echo=FALSE,warning=FALSE,message=FALSE}
mapview(Fynbos_sf)
```
The next step is cropping the fire_crop and the idx_hNDVI to only the regions of the Fynbos_sf.
```{r warning=FALSE}
fire_fynbos = st_crop(fire_crop, Fynbos_sf, crop = TRUE)

hNDVI_fynbos = st_crop(idx_hNDVI, Fynbos_sf, crop = TRUE)
```
Because the two stars objects fire_fynbos and hNDVI_fynbos have different spatial resolutions, the two objects can't be merged. This leaves conducting a spatial join of the two objects.
```{r}
fire_fynbos_sf = st_as_sf(fire_fynbos, as_points = FALSE, na.rm = TRUE)
hndvi_fynbos_sf = st_as_sf(hNDVI_fynbos, as_points = TRUE, na.rm = TRUE)
hndvi_fire = st_join(fire_fynbos_sf, hndvi_fynbos_sf, join = st_nearest_feature, left = TRUE)
```
Below is a map of the fire_fynbos_sf layer. Each polygon represents the number of days since the last fire.
```{r message=FALSE, warning=FALSE, echo=FALSE}
mapview(fire_fynbos_sf)
```
# Results

```{r message=FALSE, warning=FALSE, echo=FALSE}
ggplot(hndvi_fire, aes(x=hndvi_fire$`2022_01_17.tif`, y=hndvi_fire$PRS_L2D_STD_20220122090206_20220122090210_0001_hNDVI.tif)) +
  geom_line()
```
The graph above shows the variation of the hyperspectral NDVI over time. The time scale is the number of days since the last fire. For this region the longest time since fire came out to just over 7000 days from the collection of the data.
Based on the results of the graph above there are several different things going on with the NDVI. The first is that there were far fewer points drawn from for the longer time since fire data. Once you get over 4000 days since fire there are three primary groupings. Under 4000 days without fire there are eight groupings of data points.This is likely due to the common occurrence of wildfire in the area. This leads to less time since fire and more data points to collect from in these areas.
Looking at the two groupings of data on the graph the above 4000 days have much more consistent NDVI values than the below 4000 days. This is likely due to several factors. The first being that the below 4000 days group is subject to pioneer species that move in soon after the fire. These species are relatively short lived, but they likely crowd out what would be the more established species that take a longer time to propagate. The other factor is the location of the data point. Other influences in the area may also play an effect on the NDVI value.

# Conclusions

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.
Based on the results of this study there are two general groupings of data that can be talked about. The first is the below 4000 days since fire group. This group is defined by the large number of data points that come out of it, and the relatively inconsistent NDVI scores. The second group is the above 4000 days group. This group has fewer points, but more consistent NDVI values. This difference is likely due to the difference in plant species that inhabit the spaces during different times since fire.
There are several things that need to be further investigated with this data, and that will be covered in future research. The first is that human activity in the region needs to be accounted for. Human activity will interfere with reflectance values, and that has the possibility of skewing the data. It was not covered during this study due to size of data and time. Another factor that should be looked at is just the overall reflectance of the different time since fire areas. This would give a better explanation of what is going on in each area, and identifying things like human activity or bare soil would be easier to see and determine. Overall this is a rich data set to work with, so there are plenty of things to investigate.

# References

Busetto, L., Ranghetti, L. (2020) prismaread: A tool for facilitating access and analysis of PRISMA L1/L2 hyperspectral imagery v1.0.0, URL: https://irea-cnr-mi.github.io/prismaread/, doi: https://doi.org/10.5281/zenodo.4019081
