---
title: "Hyperspectral analysis in the Greater Cape Floristic Region of South Africa"
author: Willow Flood
date: Fall 2022 Semester
---

# Introduction

[~ 200 words]
The problem being addressed in this project is looking at the different ecosystems in the Greater Cape Floristic Region and how the reflectance is affected by wild fires. This will be done by comparing the variety of ecosystems to a time since fire layer and hyperspectral satellite data to collect the reflectance data. Background information   

# Materials and methods

[~ 200 words]

Narrative: Clear narrative description of the data sources and methods. Includes data from at least two sources that were integrated / merged in R.

Data: The underlying data are publicly accessible via the web and downloaded/accessed within the Rmd script. If you want to use your own data, you must make it available on a website (e.g. Figshare) so that others are able to re-run your code.

These are the libraries that were used in this project. (need to edit this)
```{r message=FALSE}
library(ggplot2)
library(dplyr)
library(raster)
library(rgdal)
library(plyr)
library(rhdf5)
library(hsdar)
library(piggyback)
```

These are the datasets that were used for this project.
```{r}

```

Code:

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```


```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
